Constructing Statistical Deformable Models with Shape Flow

### 2015-04-16 contents/introduction.tex
Edited Intro

We propose framework to handle training data with inconsistent annotation, build dense shape model and introduce effective curve annotation methods. In our experiment, we show building dense models trained with inconsistent annotation and trained with curve annotation improves performance of discriminative models trained on carefully annotated data.

Uses Introductions from Nondas' paper "Automatic Construction of Deformable Models In-The-Wild" as place holder.

In this paper, we deal with the problem of automatically
constructing a robust deformable model using (1) a simple
bounding box object detector and (2) a shape by means of
a Point Distribution Model (PDM). The detector can be as
simple as the Viola-Jones object detector \cite{?}
3 which returns
only a bounding box of a detected object. Such detectors
are widely employed in commercial products (e.g.
even the cheapest digital camera has a robust face detector).
Other detectors that can be used are efficient subwindow
search \cite{?}. The
annotations that are needed to train the object detector can
be acquired very quickly, since only a bounding box containing the object is required. Specifically, after selecting
the images that are going to be used, the annotation procedure
takes a couple of seconds per image. The statistical
shape model can be created by using only 40-50 shape examples,
which can be produced by either drawing possible
shape variations of the 2D shape of the object or projecting
3D CAD model instances of the object on the 2D camera
plane (such an example is shown in \cite{?} for cars). Even
the annotation of the shape examples is not a time consuming
task, due to their small number. Furthermore, there are
unsupervised techniques to learn the shape prior (model) directly
from images \cite{?}.

@@

The two most closely related works to the proposed
method are the automatic construction of Active Appearance
Models (AAMs) in \cite{?} and the so-called RASL
methodology in \cite{?} for person-specific face alignment.
There are two main differences between our framework
and \cite{?}. (1) We use a predefined statistical shape model
instead of trying to find both the shape and appearance models.
We believe that with the current available optimization
techniques, it is extremely difficult to simultaneously optimize
for both the texture and shape parameters. (2) We
employ the robust component analysis of \cite{?} for the appearance
which deals with outliers. Thus, even though
our method is similar in concept to \cite{?}, these two differences
make the problem feasible to solve. In particular, the
methodology in \cite{?} fails to create a generic model even in
controlled recording conditions, due to extremely high dimensionality
of the parameters to be found and to the sensitivity
of the subspace method to outliers. This was probably
one of the reasons why the authors demonstrate very
limited and only person-specific experiments. Furthermore,
our methodology bypasses some of the limitations of \cite{?},
which requires the presence of only one low-rank subspace,
hence it has been shown to work only for the case of congealing
images of a single person. Finally, we argue that
in order for an automatically constructed AAM methodology
to be robust to both within-class and out-of-class outliers4,
which cannot be avoided in totally unsupervised settings,
statistical component analysis techniques should be
employed \cite{?}. We summarize our contributions as follows:

We present experimental results on the task of face alignment.
We choose this application because there exist many
in-the-wild facial databases with a large number of images
and annotated landmarks, hence solid quantitative evaluations
can be performed.

>>>>

We propose a framework to (1) construct SDMs using training data with inconsistent annotation, (2) introduce highly effective curve annotation which tremendously reduces manual work overload but maintains same performance level against state-of-the-art SDMs and (3) build dense correspondent SDMs with shape flow.

% Back Ground
The two most closely related works to the proposed
method are the automatic construction of Active Appearance
Models (AAMs) in \cite{?} and the so-called RASL
methodology in \cite{?} for person-specific face alignment.
There are two main differences between our framework
and \cite{?}. (1) We use a predefined statistical shape model
instead of trying to find both the shape and appearance models.
We believe that with the current available optimization
techniques, it is extremely difficult to simultaneously optimize
for both the texture and shape parameters. (2) We
employ the robust component analysis of \cite{?} for the appearance
which deals with outliers. Thus, even though
our method is similar in concept to \cite{?}, these two differences
make the problem feasible to solve. In particular, the
methodology in \cite{?} fails to create a generic model even in
controlled recording conditions, due to extremely high dimensionality
of the parameters to be found and to the sensitivity
of the subspace method to outliers. This was probably
one of the reasons why the authors demonstrate very
limited and only person-specific experiments. Furthermore,
our methodology bypasses some of the limitations of \cite{?},
which requires the presence of only one low-rank subspace,
hence it has been shown to work only for the case of congealing
images of a single person. Finally, we argue that
in order for an automatically constructed AAM methodology
to be robust to both within-class and out-of-class outliers4,
which cannot be avoided in totally unsupervised setting
statistical component analysis techniques should be
employed \cite{?}. We summarize our contributions as follows:

% contribution summerize

In our experiment, we show that building robust dense models trained with inconsistent annotation and trained with curve annotation improves performance of discriminative models trained on carefully annotated data including ears, faces, bottles, (cars ?) and bodies. Evaluation methods and choices are described in details.

Uses Introductions from Nondas' paper "Automatic Construction of Deformable Models In-The-Wild" as place holder below.

The detector can be as
simple as the Viola-Jones object detector \cite{?}
3 which returns
only a bounding box of a detected object. Such detectors
are widely employed in commercial products (e.g.
even the cheapest digital camera has a robust face detector).
Other detectors that can be used are efficient subwindow
search \cite{?}.

The
annotations that are needed to train the object detector can
be acquired very quickly, since only a bounding box containing the object is required. Specifically, after selecting
the images that are going to be used, the annotation procedure
takes a couple of seconds per image. The statistical
shape model can be created by using only 40-50 shape examples,
which can be produced by either drawing possible
shape variations of the 2D shape of the object or projecting
3D CAD model instances of the object on the 2D camera
plane (such an example is shown in \cite{?} for cars). Even
the annotation of the shape examples is not a time consuming
task, due to their small number. Furthermore, there are
unsupervised techniques to learn the shape prior (model) directly
from images \cite{?}.

@@

% The two most closely related works to the proposed
% method are the automatic construction of Active Appearance
% Models (AAMs) in \cite{?} and the so-called RASL
% methodology in \cite{?} for person-specific face alignment.
% There are two main differences between our framework
% and \cite{?}. (1) We use a predefined statistical shape model
% instead of trying to find both the shape and appearance models.
% We believe that with the current available optimization
% techniques, it is extremely difficult to simultaneously optimize
% for both the texture and shape parameters. (2) We
% employ the robust component analysis of \cite{?} for the appearance
% which deals with outliers. Thus, even though
% our method is similar in concept to \cite{?}, these two differences
% make the problem feasible to solve. In particular, the
% methodology in \cite{?} fails to create a generic model even in
% controlled recording conditions, due to extremely high dimensionality
% of the parameters to be found and to the sensitivity
% of the subspace method to outliers. This was probably
% one of the reasons why the authors demonstrate very
% limited and only person-specific experiments. Furthermore,
% our methodology bypasses some of the limitations of \cite{?},
% which requires the presence of only one low-rank subspace,
% hence it has been shown to work only for the case of congealing
% images of a single person. Finally, we argue that
% in order for an automatically constructed AAM methodolo
% to be robust to both within-class and out-of-class outliers4,
% which cannot be avoided in totally unsupervised setti
% statistical component analysis techniques should be
% employed \cite{?}. We summarize our contributions as follows:

% We present experimental results on the task of face alignment.
% We choose this application because there exist many
% in-the-wild facial databases with a large number of images
% and annotated landmarks, hence solid quantitative evaluations
% can be performed.
